---
title: "lab 9 part 2 lizard tails chi square"
author: "Lauren Harris"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(broom)
```

Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

## Make a subset 

- only UTST lizards
- only sites with 3 highest trapped populations: - IBPE (enclosure grassland)
- RABB (rabbit mesquite site)
- EAST (east tarbush site)
- Tail condition is recorded (W - whole, B = broken)
```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>%
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```

```{r}
tail_counts <- utst %>% 
  tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
```

## Convert to a contingency table

Use `column_to_rowname()` function
```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- tidy(lizard_x2)
```
There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r round(lizard_tidy$p.value, 2)`)
